---
title: "run_example"
date: "19/06/2021"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```
  
This vignette demonstrates how to process data, set up parameters and scenarios 
for running simulation with the DynaMICE model. Details of model structure and 
assumptions can be found in a [preprint](https://www.medrxiv.org/content/10.1101/2021.05.28.21257964v1)
 entitled '*Effect of evidence updates on key determinants of measles vaccination 
 impact: a DynaMICE modelling study in ten high-burden countries*'.
  
Below we demonstrate an example run using the model incorporated with all the 
updated assumptions for the key determinants ('full-update' scenario), for the 
vaccination strategies including MCV1, MCV2, and SIA in Pakistan and India.
  
***
#### Load relevant packages
```{r load_pkg, message=FALSE}
devtools::load_all()
#library(dynamice)
#library(data.table)
#library(foreach)
#library(ggplot2)
```
  
***
#### Modify the input datasets
In this analysis, we assume perfect timeliness and fix R0 as 16 across countries.
```{r mod_input}
# assume perfect timeliness: all receive MCV1 at 9-month
adj.timeliness = TRUE  # FALSE
if (adj.timeliness){
  data_timeliness [!is.na(age) & age < 52*(9/12), timeliness := 0]
  data_timeliness [!is.na(age) & age >= 52*(9/12), timeliness := 1]
}
# assume a fixed R0
adj.fixR0 = 16  # NA
if (!is.na(adj.fixR0)){
  data_r0 [ , r0 := adj.fixR0]
}
```
  
***
#### Set up scenarios
Three types of assumptions are included for evaluation:
1. Contact matrices (POLYMOD-GBR, synthetic, proportional, uniform)
2. Age trend for vaccine efficacy (step, linear functions)
3. Proportion of SIA in zero-dose children (random, logistic function)
Some assumptions are modified in fortran and others in R. 
```{r setup_scn}
fortran_names <- c("vaccine2019_sia_adjBeta_mcv2.exe",   # most updated version
                   "vaccine2019_sia_adjBeta_mcv2_nodpdSIA.exe",
                   "vaccine2019_sia_adjBeta_mcv2_stepVE.exe",
                   "vaccine2019_sia_adjBeta_mcv2_nodpdSIA_stepVE.exe",
                   "vaccine2019_sia_adjBeta_mcv2_stepVE_SIAsa.exe"  # sensitivity analysis for SIA delivery
                   )
					  
contact_mats <- c("syn", "polymod", "prpmix", "unimix")

main <- list (
  scenarios    = c("Base",          # (1) Base: step change VE + random SIA delivery to zero-dose children
                   "Efficacy",      # (2) Base + linear VE
                   "SIAdeliver",    # (3) Base + non-random SIA delivery to zero-dose children
                   "ContactSyn",    # (4) Base + synthetic contact patterns
                   "ContactPrp",    # (5) Base + proportional mixing
                   "ContactUni",    # (6) Base + uniform mixing
                   "Update",        # (7) Update: linear VE + synthetic matrix + non-random SIA delivery
                   "SIAdeliverSA"   # (8) Base + sensitivity analysis for non-random SIA delivery
                  ),
  fortran_model = fortran_names [c(4, 2, 3, 4, 4, 4, 1, 5)],
  contact_mat   = contact_mats  [c(2, 2, 2, 1, 3, 4, 1, 2)],
  step_ve       =                c(T, F, T, T, T, T, F, T)
)
```
  
***
#### Compile fortran codes
The simulation model was constructed using fortran, to take the advantage of its 
fast computation speed. The files containing fortran codes (*.F95) in the 
`model/fortran` folder need to be compiled into executable files (*.exe) with the
 same name and then saved in the `model/compiled` folder.
  
**Mac OS**  
1. Install [Homebrew](https://brew.sh/), an open-source software package management system
for Mac OS or Linux  
2. Install *GNU compiler collection (GCC)* via terminal: `brew install gcc`  
3. To test that .F95 files can be compiled using this newly installed software, it 
may be worthwhile to [run a 'hello world' program](https://fortran-lang.org/learn/building_programs/compiling_source).  
  
**Windows**  
1. Install [cygwin](https://www.cygwin.com/), a runtime environment that provides POSIX 
compatibility to Windows  
2. Install the *GCC* package if not already built in  
  
With the installed compiler, use *gfortran* (part of *GCC*) to compile each file in
the `model/fortran` folder via terminal or cygwin terminal:
`gfortran -o ../compiled/vaccine2019_sia_adjBeta_mcv2.exe vaccine2019_sia_adjBeta_mcv2.F95 -fcheck=bounds -ffree-line-length-0`  
  
***
#### Specify the scenario for evaluation
```{r select_scn}
imain <- 7
print (paste0 ("Evaluation started: ", main$scenarios[imain]))
```
  
***
#### Assign file paths and global variables
```{r assign_var}
var <- list (
  # vaccine coverage
  vaccine_coverage_folder           = "vac_coverage_top10/",
  coverage_prefix                   = "coverage",
  touchstone                        = "_",
  antigen                           = NULL,
  vaccine_coverage_subfolder        = "scenarios/",

  # disease burden
  central_burden_estimate_folder    = "central_burden_estimate/",
  stochastic_burden_estimate_folder = "stochastic_burden_estimate/",

  # diagnostic plots folder
  plot_folder                       = "plots/",

  # modelling group name
  group_name                        = NULL,
  
  # log file name
  log_name                          = "test_log",

  # countries - specify iso3 codes to analyse only these countries
  #             or set it to "all" to analyse all included countries
  countries                         = c("PAK"),#,"IND"

  cluster_cores                     = 1,    # number of cores 
  psa                               = 0     # psa runs; 0 for single central run
)

# for central run: set number of runs to 0 (var$psa = 0)
# for stochastic runs: set number of runs to greater than 0 (eg: var$psa = 200)
# burden_estimate_folder (central or stochastic)
if (var$psa == 0) {
  burden_estimate_folder <- var$central_burden_estimate_folder
} else {
  burden_estimate_folder <- var$stochastic_burden_estimate_folder
}

# folders for burden estimate results
dir.create (file.path (paste0 (getwd(), "/", burden_estimate_folder, "Wolfson/")), recursive = T)
dir.create (file.path (paste0 (getwd(), "/", burden_estimate_folder, "Portnoy/")), recursive = T)
```
  
***
#### Set up vaccination strategies
```{r setup_vac}
vaccine_strategies <- c("no-vacc",    # (1) no vaccination (set vaccination and using_sia to 0)
                        "mcv1",       # (2) MCV1 only
                        "mcv2",       # (3) MCV1 & MCV2
                        "sia-conti",  # (4) MCV1 & MCV2 and SIAs
                        "sia-stop"    # (5) MCV1 & MCV2 and SIAs (stop SIAs after 2019)
                        )

# corresponding vaccination strategies used for Portnoy's CFR
portnoy_scenarios <- c("no-vaccination",    # (1) no vaccination (set vaccination and using_sia to 0)
                       "mcv1-default",      # (2) MCV1 only
                       "mcv2-default",      # (3) MCV1 & MCV2
                       "campaign-default",  # (4) MCV1 & MCV2 and SIAs
                       "campaign-default"   # (5) MCV1 & MCV2 and SIAs (stop SIAs after 2019)
                       )

# set SIAs and vaccination parameters for each scenario to minimize errors for running
set_sia         <- c (0, 0, 0, 1, 1)
set_vaccination <- c (0, 1, 2, 2, 2)
```

***
#### Prepare coverage files
Pre-processing of vaccine coverage data are needed and additional files are 
created for the model input, including:  
1. Adding the coverage file for 'no-vaccination' strategy,  
2. Adding the coverage file for 'stop-future-campaign' strategy,  
3. Separating coverage for routine and campaign immunisations into two files for
each vaccination strategy.  
```{r setup_vaccov}
update_coverage_inputs = FALSE
if (update_coverage_inputs) {
  # (1) create vaccine coverage file (0% coverage) for no vaccination scenario
  create_no_vaccination_coverage_file (
    no_vaccination_coverage_file = paste0 (var$vaccine_coverage_folder,
                                           "coverage_", scenarios[1],".csv"),
    vaccination_coverage_file    = paste0 (var$vaccine_coverage_folder,
                                           "coverage_", scenarios[2],".csv")
  )

  # (2) create a vaccine coverage file for no future SIA scenario
  continuous_sia_coverage <- fread (paste0(var$vaccine_coverage_folder,
                                           "coverage_", scenarios[4],".csv"))
  stop_sia_coverage <- continuous_sia_coverage [!(year > 2019 & activity_type == "campaign")]
  fwrite (stop_sia_coverage, file = paste0(var$vaccine_coverage_folder,
                                           "coverage_", scenarios[5],".csv"))

  # (3) generate 2 vaccine coverage files per scenario for routine and SIA
  for (index in 1:5) {
    create_vaccine_coverage_routine_sia (
      vaccine_coverage_folder    = var$vaccine_coverage_folder,
      vaccine_coverage_subfolder = var$vaccine_coverage_subfolder,
      coverage_prefix            = var$coverage_prefix,
      touchstone                 = var$touchstone,
      antigen                    = var$antigen,
      scenario_name              = scenarios [index],
      rev_cov                    = FALSE # coverage data have been pre-processed
      )
  }
}
```

***
#### Create samples for probabilistic sensitivity analysis
For PSA, a latin hyper cube sample is generated for the 3 parameters: case 
fatality risk, intercept and slope for the age-dependent function for first-dose
vaccine efficacy.  
```{r create_psa}
if (var$psa > 0) {
  psadat <- create_PSA_data (psa             = var$psa,
                             seed_state      = 1,
                             psadat_filename = "psa_variables.csv")
}
```
  
***
#### Get reulsts from a selected vaccination strategy
The DynaMICE model produces case estimates, which are then used to calculate 
deaths and disability-adjusted life years by Wolfson's and Portnoy's CFRs. 
Results by country, calendar year, and single age strata are saved as csv files 
under the previously specified route.
  
```{r get_results}
# selected scenario
index <- 4
scenario_name  <- vaccine_strategies [index]
print (scenario_name)
scenario_number <- sprintf ("scenario%02d", index)

# run model and estimate cases
burden_estimate_file <- runScenario (
  vaccine_coverage_folder    = var$vaccine_coverage_folder,
  vaccine_coverage_subfolder = var$vaccine_coverage_subfolder,
  coverage_prefix            = var$coverage_prefix,
  antigen                    = var$antigen,
  scenario_name              = scenario_name,
  save_scenario              = scenario_number,
  burden_estimate_folder     = burden_estimate_folder,
  group_name                 = var$group_name,
  log_name                   = var$log_name,
  countries                  = var$countries,
  cluster_cores              = var$cluster_cores,
  psa                        = var$psa,
  vaccination                = set_vaccination [index],
  using_sia                  = set_sia         [index],
  measles_model              = main$fortran_model [imain], 
  debug_model                = FALSE,
  contact_mat                = main$contact_mat [imain],
  step_ve                    = main$step_ve     [imain],
  sim_years                  = 1980:2100
  )

# estimate deaths and DALYs by Wolfson and Portnoy's methods
estimate_deaths_dalys (cfr_option             = "Wolfson",
                       burden_estimate_file   = burden_estimate_file,
                       burden_estimate_folder = burden_estimate_folder,
                       psa                    = var$psa)

estimate_deaths_dalys (cfr_option             = "Portnoy",
                       burden_estimate_file   = burden_estimate_file,
                       burden_estimate_folder = burden_estimate_folder,
                       vimc_scenario          = portnoy_scenarios[index],
                       portnoy_scenario       = "s6",  # constant CFR after 2018
                       psa                    = var$psa)
```
  
***
#### Get diagnostic plots
For central estimates, simple diagnostics plots for vaccine coverage and disease
 burden (cases, deaths, DALYs) are generated and then saved in a pdf file for 
 each country.
```{r plot_diagnose}
if (var$psa == 0) {

  dir.create (file.path (paste0 (getwd(), "/", var$plot_folder)), recursive = T)

  diagnostic_plots (
    vaccine_coverage_folder    = var$vaccine_coverage_folder,
    coverage_prefix            = var$coverage_prefix,
    touchstone                 = var$touchstone,
    antigen                    = var$antigen,
    scenarios                  = vaccine_strategies[index],
    base_scenario              = "no-vaccination",
    burden_estimate_folder     = var$central_burden_estimate_folder,
    plot_folder                = var$plot_folder,
    group_name                 = var$group_name,
    countries                  = var$countries,
    cfr_options                = c("Wolfson", "Portnoy"),
    psa                        = var$psa,
    start_year                 = 2000,
    end_year                   = 2050,
    compare_plots              = FALSE
  )
}
```
